Seg_sites_rarefaction

### 2024-04-22 scripts/seg_sites_rarefaction.Rmd
files pre-publication for movement to christeLab repo

```{r cluster_MCMCs, eval=FALSE}
# set.seed(seed + q) # reset seed for cluster

@@

# import and condense simulation results
```{r}
res_files <- list.files("../data/", "\\.RDS", full.names = TRUE)
res_files <- lapply(res_files, readRDS)
ks <- unlist(purrr::map(res_files, "k"))
n_iters <- unlist(purrr::map(res_files, "iters"))

@@

```{r Figure1}
prep_f1 <- function(pseg1, pseg2, mpseg1, mpseg2, n1, n2){
  #========plot the emperical CIs around every loci vs the predicted value==============
  # plotting the emperical CIs vs the estiamted mean
  pseg <- rbind(data.frame(pop = paste0("N = ", n1/2), p = pseg1, snp = 1:nal, af = allele_frequencies),
                data.frame(pop = paste0("N = ", n2/2), p = pseg2, snp = 1:nal, af = allele_frequencies))

@@

# Figure 2: var vs difference between estimate and mean of simulations for mcmc and k
Note that the labels are flipped here, this produces figure 2

@@

  ps_prep$K <- tk
  ps_prep$iters <- titers
  ps_prep$sim_num <- i

  ns_prep$all_mcmc_sums$K <- tk
  ns_prep$all_mcmc_sums$iters <- titers
  ns_prep$all_mcmc_sums$i <- i
  ns_prep$all_mcmc_sums$ENs <- ns_prep$sum_tab$nseg[match(ns_prep$all_mcmc_sums$pop,
                                                          ns_prep$sum_tab$pop)]

@@

ps_all_m <- ps_all[source == "Mathematical",]
ps_all_s <- ps_all[source == "Simulation",]
ps_all_merge <- merge(ps_all_m[,-c("source")], 
                      ps_all_s[,-c("source")], 
                      by = c("af", "pop", "snp", "sim_num", "lower", "upper", "K", "iters"), 
                      suffixes = c(".mathematical", ".simulated"))
ps_all_merge[,diff := p.mathematical - p.simulated]

@@

save_plot("FigureS1.pdf", FS1, base_height = 10, base_width = 15)

@@

save_plot("Figure2.pdf", f3, base_height = 11, base_width = 15)
```
# Figure 4 Nseg vs parameters

@@

save_plot("FigureS2.pdf", FS2, base_height = 10, base_width = 15)

@@

d <- calc_seg_sites(d, "pop", g = 0) # set gamma to the smallest N per loci
seg_res_mon <- get.snpR.stats(d, "pop", "seg_sites") # follows biogeographic history
```

@@

ggplot(mon_rare_res, aes(x = pop, y = Ns)) +
  geom_boxplot() +
  geom_point(alpha = .2, position = position_jitter(height = 0, width = .2)) +
  geom_point(data = seg_res_mon$weighted.means, aes(x = subfacet, y = seg_sites), color = "red") +
  theme_bw() +
  ylab(bquote(italic(N[s]))) +
  xlab("Population")

>>>>

```{r cluster_MCMCs}
# set.seed(seed + q) # reset seed for cluster

@@

# import and condense simulation results
```{r}
res_files <- list.files("../data/", "mcmc.+\\.RDS", full.names = TRUE)
res_files <- lapply(res_files, readRDS)
ks <- unlist(purrr::map(res_files, "k"))
n_iters <- unlist(purrr::map(res_files, "iters"))

@@

```{r Figure1}
prep_f1 <- function(pseg1, pseg2, mpseg1, mpseg2, n1, n2){
  #========plot the empirical CIs around every loci vs the predicted value==============
  # plotting the empirical CIs vs the estimated mean
  pseg <- rbind(data.frame(pop = paste0("N = ", n1/2), p = pseg1, snp = 1:nal, af = allele_frequencies),
                data.frame(pop = paste0("N = ", n2/2), p = pseg2, snp = 1:nal, af = allele_frequencies))

@@

# Figure 2: var vs difference between estimate and mean of simulations for mcmc and k; Figure S1
Note that the labels are flipped here, this produces figure 2

@@

  # per snp
  ps_prep$K <- tk
  ps_prep$iters <- titers
  ps_prep$sim_num <- i

  ## counts seg per snp
  cn1 <- res_files[[i]]$seg1
  cn1 <- rowSums(cn1)
  cn2 <- res_files[[i]]$seg2
  cn2 <- rowSums(cn2)

  cn <- rbind(data.frame(pop = paste0("N = ", n1/2), af = allele_frequencies, n_obs_seg = cn1),
              data.frame(pop = paste0("N = ", n2/2), af = allele_frequencies, n_obs_seg = cn2))
  ps_prep <- merge(ps_prep, cn, by = c("pop", "af"), sort = FALSE)

  # total
  ns_prep$all_mcmc_sums$K <- tk
  ns_prep$all_mcmc_sums$iters <- titers
  ns_prep$all_mcmc_sums$i <- i
  ns_prep$all_mcmc_sums$ENs <- ns_prep$sum_tab$nseg[match(ns_prep$all_mcmc_sums$pop,
                                                          ns_prep$sum_tab$pop)]

@@

ps_all_m <- ps_all[source == "Mathematical",] # note: all of the CIs in this refer to: is the mathematical expectation in the observed CI.
ps_all_s <- ps_all[source == "Simulation",]
ps_all_merge <- merge(ps_all_m[,-c("source")], 
                      ps_all_s[,-c("source")], 
                      by = c("af", "pop", "snp", "sim_num", "lower", "upper", "K", "iters"), 
                      suffixes = c(".mathematical", ".simulated"))
ps_all_merge[,diff := p.mathematical - p.simulated]

# note: this code does the reverse: is the observed number of seg sites for a locus a set of sims within the PI expected based on the expected prob
ps_all_merge[,pi_lower := qbinom(.05, iters, p.mathematical)]
ps_all_merge[,pi_upper := qbinom(.95, iters, p.mathematical)]
ps_all_merge[,in_pi := n_obs_seg.simulated >= pi_lower & n_obs_seg.simulated <= pi_upper]

@@

save_plot("FigureS1.pdf", FS1, base_height = 15, base_width = 15)
save_plot("FigureS1.jpg", FS1, base_height = 15, base_width = 15)

@@

save_plot("Figure2.pdf", f3, base_height = 11, base_width = 15)
```
# Figure 4 Nseg vs parameters; Figure S2

@@

save_plot("FigureS2.pdf", FS2, base_height = 10, base_width = 21)
save_plot("FigureS2.jpg", FS2, base_height = 10, base_width = 21)

@@

d <- calc_seg_sites(d, "pop", g = 0) # set gamma to the smallest N per loci
seg_res_mon <- get.snpR.stats(d, "pop", "seg_sites") # follows biogeographic history

# save this as a VCF for the supplementary material example
format_snps(d, "vcf", outfile = "../data/example_vcf.vcf", chr = "CHROM")
fwrite(sample.meta(d), "../data/example_metadata.txt", col.names = TRUE, row.names = FALSE, sep = "\t")
```

@@

FS3 <- ggplot(mon_rare_res, aes(x = pop, y = Ns)) +
  geom_boxplot() +
  geom_point(alpha = .2, position = position_jitter(height = 0, width = .2)) +
  geom_point(data = seg_res_mon$weighted.means, aes(x = subfacet, y = seg_sites), color = "red", size = 4) +
  theme_bw() +
  ylab(bquote(italic(N[s]))) +
  xlab("Population") +
  theme(axis.text = element_text(size = t.sizes[2]),
        axis.title = element_text(size = t.sizes[1]),
        legend.title = element_text(size = t.sizes[1]))

save_plot("FigureS3.pdf", FS3, base_width = 15, base_height = 11)
save_plot("FigureS3.jpg", FS3, base_width = 15, base_height = 11)
```
